---
title: "Frogs"
author: "Ben Williams"
date: "July 30, 2015"
output: html_document
---
First I changed the data file to a .csv
Import data and re-label
```{r input}
Htest_R.input2 <- read.csv("Htest_R input2.csv")

# change file name
dat <- Htest_R.input2
# changes column names
names(dat) <- c("pond", "id", "abnormal", "hetero", "road", "oil")

dat$Road <- factor(dat$road)
dat$Oil <- factor(dat$oil)
dat$Abnormal <- factor(dat$abnormal)
dat$Pond <- factor(dat$pond)


```
#Plots

Plot the individual factors by level (0,1). Did this for oil, road, and oil/road interaction - doesn't take into account repeated samples from same ponds.
```{r plot}
library(ggplot2)
theme_set(theme_bw(base_size=12)+theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank()))

ggplot(dat, aes(hetero, abnormal, group=Oil, color=Oil, fill=Oil))+
   stat_smooth(family='binomial', method='gam', formula=y~s(x, k=4), alpha=.3)+ylim(0,1)+
   theme(legend.justification=c(1,0), legend.position=c(.9,.7))
```

The above figure shows no real difference between close to or far from oil producing locations - there is a slightly negative relationship between heterozygosity and the presence of abnormalities, away from oil patches.

```{r plot 2}
ggplot(dat, aes(hetero, abnormal, group=Road, color=Road, fill=Road))+
   stat_smooth(family='binomial', method='gam', formula=y~s(x, k=4), alpha=.3)+ylim(0,1)+
   theme(legend.justification=c(1,0), legend.position=c(.9,.7))
```

The above figure shows a somewhat signiificant relationship close to or far from a road with frogs further from roads having higher levels of abnormalities.

```{r plot 3}
ggplot(dat, aes(hetero, abnormal, group=interaction(Road, Oil), color=interaction(Road,Oil), fill=interaction(Road,Oil)))+
      stat_smooth(family='binomial', method='gam', formula=y~s(x, k=4),alpha=.3)+ylim(0,1)+
   theme(legend.justification=c(1,0), legend.position=c(.9,.7))
```

The above figure shows interactions between being close to oil patch and road or far from, or a combination between the two. If a frog is not close to an oil patch or road there is a negative relationship between heterozygosity and abnormalities (though it is weak). If a frog is close to either a road or oil patch, but not both then a) all of the samples have heterozygosity greater than 0, and b) there is no evident relationship between heterozygosity and abnormalities. If a frog is both close to a road and an oil patch there is no significant relationship evident, though there appears to be a trend with higher levels of abnormalities at 0 and heterozygosity levels around 1.5.


#Models

Use a GAM for modeling random effects (could also easily use glmer in the lme4 package). 

First examine without random effects
```{r model}
library(mgcv)
m1 <- gam(Abnormal~s(hetero)+Road+Oil, data=dat, gamma=1.4, family = binomial)
summary(m1)
```

This provides the same general evaluation as the above figures do: There is a relationship that is significant when oil and road are 0 (intercept is significant), there is a signhificant effect of road, but not of being close to an oil patch. Further there is no direct relationship between heterozygosity and abnormalities.

Same model with random effects for pond <- this is a more appropriate method than the one above...

```{r model2}
m2 <- gam(Abnormal~s(hetero)+s(Pond, bs='re')+Road+Oil, data=dat, gamma=1.4, family = binomial)
summary(m2)
```

This shows pretty much the same results, though now the road factor is only marginally significant. The locations absorb a fair bit of variability and the overall model explains more deviance, though still isn't that good as a descriptor.

Examine if either the road or oil aspects influence heterozygosity.
```{r model3}
m3 <- gam(hetero~s(Pond, bs='re')+Road+Oil, data=dat, gamma=1.4)
summary(m3)
```

This model is a poor descriptor, though it does show a significant negative association between oil patch proximity and heterozygisity, there is no significant relationship with proximity to a road.


#Wrap up

1. Looks like there is not a relationship in these data between heterozygosity and frog abnormalities,
2. There is an effect by location - would be helpful to plot out a map of abnormalities and examine spatial effects,
3. There is a negative relationship between oil patch proximity and heterozygosity.

Short answer, these data do not appear to hold a 'smoking gun' in regard to indetifying a relationship to frog abnormalities.
